\documentclass[a4paper,twocolumn]{article}

\usepackage{times}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[a4paper,margin=2cm,columnsep=1cm]{geometry}
\usepackage{authblk}
\usepackage{titlesec}
\usepackage[pdftex]{graphics}
\usepackage{mathtools}
\usepackage{enumitem}
\usepackage{amssymb}

\topmargin      0.0cm
\headheight     0.0cm
\headsep        0.0cm
\oddsidemargin  0.0cm
\evensidemargin 0.0cm
\textheight     22.86cm
\textwidth      16.51cm


\begin{document}

\graphicspath{{images/}}
\titleformat*{\section}{\normalsize\bfseries\filcenter}
\titleformat*{\subsection}{\normalsize\bfseries}
\renewcommand{\figurename}{\small Figure}
\newcommand{\figureref}[1]{Fig. (\ref{fig:#1})}
\newcommand{\equationref}[1]{Eq. (\ref{eq:#1})}
\newcommand{\bigsum}{\displaystyle\sum}

\title{
    \textbf{Final Term Paper}\\
    \textit{Partial Report}}
\author{
    \textbf{Eduardo Martins Barros de Albuquerque Tenório} (\textit{student})\\
    \textbf{Tsang Ing Ren} (\textit{advisor})\\
    \small{\texttt{\{embat,tir\}@cin.ufpe.br}}
}
\affil{\large
    Centro de Informática,\\
    Universidade Federal de Pernambuco
}
\date{\today}

\maketitle


This document contains all partial reports from the project, organized by date (oldest on top). Each report contains a brief description of the work performed, delays, changes on schedule and a prevision of what may be on next.


\section*{Reports}
\label{ch:reports}

\subsection*{2014-11-25}

As the proposal was delivered at the end of October, this report just concerns about November.

The schedule is delayed due to activities from other courses. Only half of what was accorded is done (as marked on the list below). From the unmarked items, the UBM-GMM is programmed to be finished at the end of November (weekend) and the adapted speaker GMM and FCM implementation to be closed on the first week of December.

The references searched and read account 1 PhD thesis about GMM (Douglas A. Reynolds), 22 papers about MFCC extraction, GMM-UBM, Adapted GMM and FCM, and 1 tutorial about UBM.

\begin{itemize}[noitemsep]
    \item Literature research. $\text{\rlap{$\checkmark$}}$
    \item MFCCs extraction. $\text{\rlap{$\checkmark$}}$
    \item Corpus read. $\text{\rlap{$\checkmark$}}$
    \item Simple speaker GMM. $\text{\rlap{$\checkmark$}}$
    \item UBM-GMM.
    \item Adapted speaker GMM.
    \item FCM.
\end{itemize}

The schedule for December is changed to:

\begin{itemize}[noitemsep]
    \item Adapted speaker GMM development and testing.
    \item FCM development and test.
\end{itemize}

\end{document}